Nested Sampling with Multiple Objective Functions

### 2018-06-03 doc/ms.tex
putting citations in

% Author Orchid ID: enter ID or remove command
\newcommand{\xx}{\boldsymbol{\theta}}
\newcommand{\dx}{d\boldsymbol{\theta}}
\newcommand{\data}{D}
\newcommand{\II}{I}

@@

Nested Sampling \citep[NS][]{skilling} is general
Monte Carlo algorithm that can solve a wide range of problems in Bayesian
computation and statistical mechanics
\citep{2009arXiv0906.3544P, 2014PhRvE..89b2302P, 2015arXiv150303404B}.
Its key strength its ability to cope with phase transitions and other
features which cause
problems for many other methods.
In a Bayesian inference problem with unknown parameters denoted collectively
by a vector $\xx$, the
posterior distribution for the parameters given a data proposition $\data$ and
a `prior information' proposition $\II$ is:
\begin{eqnarray}
p(\xx | \data, \II) &=&
\frac{p(\xx | \II)p(\data | \xx, \II)}{p(\data | \II)}\\
&=& \frac{\pi(\xx)L(\xx)}{Z}
\end{eqnarray}
where $\pi(\xx)$ is the prior distribution, $L(\xx)$ is the likelihood
function, and $Z$ is the normalising constant, known as the
`marginal likelihood' or `evidence':
\begin{eqnarray}
Z &=& \int \pi(\xx) L(\xx) \, \dx.\label{eqn:evidence}
\end{eqnarray}

@@

The main goal of NS is to compute $Z$, but it can also be used to make
Monte Carlo approximations of the posterior distribution, and to
calculate the `information',
or Kullback-Leibler divergence from the prior to the posterior:
\begin{eqnarray}
H &=& \int p(\xx | \data, \II) \log
\left(\frac{p(\xx | \data, \II)}{p(\xx | \II)}\right) \, d\xx \\
&=& \int \frac{\pi(\xx) L(\xx)}{Z} \log
\left(\frac{L(\xx)}{Z}\right) \, d\xx.
\end{eqnarray}
$H$ quantifies how compressed the posterior distribution is with
respect to the prior. For example, $H = 100$ nats (the units obtained when
the logarithm is the natural logarithm) implies, loosely speaking,
that the posterior occupies a fraction $e^{-100}$ of the prior mass.
$H$ can be interpreted quite literally as a measure of how much
the specific data $\data$ resolved the question ``what is the value
of $\xx$, precisely?'' \citep{knuth, van_erp}.

@@

We seek an algorithm for computing the partition function
$Z(\beta_1, \beta_2)$ for a range of values of the $\beta$s, from a single
run. To be a variant of Nested Sampling, the algorithm should satisfy
the following properties:
\begin{enumerate}
\item It should begin with $N$ points drawn from the prior $\pi(\xx)$.
\item It should seek to explore regions where the values of
$L_1(\xx)$ and $L_2(\xx)$ are higher than the prior $\pi(L_1, L_2)$
would typically imply.
\item The algorithm should consider a sequence of probability
distributions proportional to $\pi$, but restricted to smaller and smaller
domains for which the enclosed prior mass shrinks by a (roughly) constant and
known factor at each iteration.
\item The algorithm should be invariant to monotonic transformations of
$L_1$ and $L_2$, i.e. it should only depend on rankings of $L_1$ and $L_2$
values and not the values themselves.
\end{enumerate}
The standard Nested Sampling algorithm has these properties but for a
single objective function. The same is true of variants such as
Diffusive Nested Sampling \citep{dnest}.

@@

\bibliographystyle{mdpi}
\bibliography{references/references}

>>>>

% Author Orchid ID: enter ID or remove command
\newcommand{\xx}{\boldsymbol{\theta}}
\newcommand{\dx}{d\boldsymbol{\theta}}
\newcommand{\data}{D}
\newcommand{\II}{I}
\newcommand{\mnras}{Monthly Notices of the Royal Astronomical Society}

@@

Nested Sampling \citep[NS][]{skilling2006nested} is general
Monte Carlo algorithm that can solve a wide range of problems in Bayesian
computation and statistical mechanics
\citep[e.g.][]{2009arXiv0906.3544P, 2014PhRvE..89b2302P, 2015arXiv150303404B}.
Its key strength its ability to cope with phase transitions and other
features which cause
problems for many other methods.
In a Bayesian inference problem with unknown parameters denoted collectively
by a vector $\xx$, the
posterior distribution for the parameters given a data proposition $\data$ and
a `prior information' proposition $\II$ is:
\begin{eqnarray}
p(\xx | \data, \II) &=&
\frac{p(\xx | \II)p(\data | \xx, \II)}{p(\data | \II)}\\
&=& \frac{\pi(\xx)L(\xx)}{Z}
\end{eqnarray}
where $\pi(\xx)$ is the prior distribution, $L(\xx)$ is the likelihood
function, and $Z$ is the normalising constant, known as the
`marginal likelihood' or `evidence':
\begin{eqnarray}
Z &=& \int \pi(\xx) L(\xx) \, \dx.\label{eqn:evidence}
\end{eqnarray}

@@

The main goal of NS is to compute $Z$, but it can also be used to make
Monte Carlo approximations of the posterior distribution, and to
calculate the `information',
or Kullback-Leibler divergence from the prior to the posterior:
\begin{eqnarray}
H &=& \int p(\xx | \data, \II) \log
\left(\frac{p(\xx | \data, \II)}{p(\xx | \II)}\right) \, d\xx \\
&=& \int \frac{\pi(\xx) L(\xx)}{Z} \log
\left(\frac{L(\xx)}{Z}\right) \, d\xx.
\end{eqnarray}
$H$ quantifies how compressed the posterior distribution is with
respect to the prior. For example, $H = 100$ nats (the units obtained when
the logarithm is the natural logarithm) implies, loosely speaking,
that the posterior occupies a fraction $e^{-100}$ of the prior mass.
$H$ can be interpreted quite literally as a measure of how much
the specific data $\data$ resolved the question ``what is the value
of $\xx$, precisely?'' \citep{knuth_questions, van_erp}.

@@

We seek an algorithm for computing the partition function
$Z(\beta_1, \beta_2)$ for a range of values of the $\beta$s, from a single
run. To be a variant of Nested Sampling, the algorithm should satisfy
the following properties:
\begin{enumerate}
\item It should begin with $N$ points drawn from the prior $\pi(\xx)$.
\item It should seek to explore regions where the values of
$L_1(\xx)$ and $L_2(\xx)$ are higher than the prior $\pi(L_1, L_2)$
would typically imply.
\item The algorithm should consider a sequence of probability
distributions proportional to $\pi$, but restricted to smaller and smaller
domains for which the enclosed prior mass shrinks by a (roughly) constant and
known factor at each iteration.
\item The algorithm should be invariant to monotonic transformations of
$L_1$ and $L_2$, i.e. it should only depend on rankings of $L_1$ and $L_2$
values and not the values themselves.
\end{enumerate}
The standard Nested Sampling algorithm has these properties but for a
single objective function. The same is true of variants such as
Diffusive Nested Sampling \citep{dns, dnest4}.

@@

%\bibliographystyle{mdpi}
\bibliography{references.bib/references}
